---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

First we load al the required packages we are going to use 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Loading and preprocessing the data

Here we read the data by using read.csv 

```{r}
data <- read.csv(unz("activity.zip","activity.csv"))
```

We inspect the data cheking the first rows

```{r}
head(data)
```

We clean the data by removing all rows where `NA` values are found

```{r}
clean_data <- data[complete.cases(data),]
```

## What is mean total number of steps taken per day?

Then we calculated the total number of steps per day, by grouping and summing the information in data

```{r}
data_per_day <- clean_data %>% group_by(date) %>%
  summarise(mean=mean(steps), sum=sum(steps))
```

We check this information by showing the histogram of the steps per day, binned in ten different ranges.
```{r}
hist(as.numeric(data_per_day[['sum']]), breaks = 10, main = "Total numbers of steps per day", xlab = "Number of steps per day")
```

Then the mean and median total number of steps per day will be
```{r}
lapply(data_per_day['sum'], mean)
lapply(data_per_day['sum'], median)
```

## What is the average daily activity pattern?

Now we check the five minute intervals of the steps taken

```{r}
data_per_5min <- clean_data %>% group_by(interval) %>%
  summarise(mean=mean(steps), sum=sum(steps))
```

Then we plot the the time series of these intervals with respect to the number of steps for all days

```{r}
ggplot(data = data_per_5min, aes(x = interval, y = mean))+ geom_line(color="#00AFBB", size=2) + labs(title = "Average daily pattern", x="5 minute interval", y= "Average number of steps")
```
Then we search for the interval with the higher number of steps

```{r}
data_per_5min$interval[which.max(data_per_5min$mean)]
```

## Imputing missing values

In the original `data` we have some `NA` values, here we count those `NA` in the steps column (As this column is the only one with `NA`)

```{r}
sum(is.na(data))
```

There are many ways to impute these data, but the most logical is to use the integer of the 5 minute average for those intervals. We impute those values here by merging the two datasets and then replacing the `steps` values with the mean where there is a `NA`, finally we order by date

```{r}
merged_data <- merge(data_per_5min, data, by = c("interval"), all.y = TRUE, sort = FALSE)
merged_data$steps <- ifelse(is.na(merged_data$steps), as.integer(merged_data$mean), merged_data$steps)
merged_data$mean <- NULL
merged_data$sum <- NULL
merged_data <- merged_data[order(merged_data$date),]
head(merged_data)
```
Then we create the same histogram plot and the mean and median as before to check that the trends are very similar

```{r}
mdata_per_day <- merged_data %>% group_by(date) %>%
  summarise(sum=sum(steps))

hist(as.numeric(mdata_per_day[['sum']]), breaks = 10, main = "Total numbers of steps per day", xlab = "Number of steps per day")
```
This is pretty similar to the one show before. 

```{r}
lapply(mdata_per_day['sum'], mean)
lapply(mdata_per_day['sum'], median)
```
The values change a bit due to the imputation, one of the impacts is that some days where there was no data, now are shoing as the average of the 5 minute intervals, so changing the data in some days change the median more drastically. 

## Are there differences in activity patterns between weekdays and weekends?


